---
title: "test1"
output: 
  md_document:
    variant: markdown_github
  
---

Loading needed libraries:
```{r}
library(FactoMineR)
library(factoextra)
library(ggpubr)
```

Use multiple correspondence analysis (MCA) to explain the variability between these
potential customers in terms of the product likes and dislikes. When you look at a plot
of the individuals one of them will really standout. You have just identified an extreme
outlier in these data! Look at the profile of ratings for this person using the View()
command in R to view of spreadsheet of these data. Can you see anything unique
about this individual? What is it? 


```{r}
prodrat = read.csv("C:/Users/fb8502oa/Desktop/DSCI 415/DSCI 415 fall/ProductRatings.csv", header=TRUE)
names(prodrat)

row.names(prodrat) = prodrat$QN
prod.mat = prodrat[,-1]
prod.mat = apply(prod.mat,2,as.factor)
```

```{r}
#doing the mca
prod.MCA = MCA(prod.mat, graph = FALSE)
summary(prod.MCA)
```

```{r}
#ploting the individual 
fviz_mca_ind(prod.MCA, col.ind = "steelblue") 
#outliyer = indv 129
```
#outliyer = indv 129

```{r}
View(prod.mat)# to see which column 129 is at
```


Remove the outlying individual in part (a) and conduct a thorough MCA of these data.
Construct plots showing individuals, variables (product ratings), and both
individuals & variables. Use the “cos2” measure of quality of representation to
color code the individual painting in at least one of your displays and use the
“contrib” measure of importance to color code the variables in at least one of
your representations. 

```{r}
#removing indv 129
newprodrat = prod.mat[-c(4),]


#mca of the new data 
newprod.MCA = MCA(newprodrat)
summary(newprod.MCA)

```

comments = 

```{r}
#ploting for the individual
fviz_mca_ind(newprod.MCA, col.ind = "cos2")

```

comments =

```{r}
#ploting for the variables
fviz_mca_var(newprod.MCA, col.var = "contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"))
```

comments = 

```{r}
#ploting for both the variables and the products
fviz_mca_biplot(newprod.MCA, col.var ="#FC4E07") #this graph is terrible so imma try ploting usinf usdv
plot(newprod.MCA)
```

comments =



Use a biplot (i.e. plot of both individuals and variables) to discuss the quadrants in
terms of the product ratings and identify groups of individuals with similar product
preferences. Discuss your findings

```{r}
biplotprodrat = prodrat[-c(4),]
biplotprodrat.prop = as.matrix(biplotprodrat / sum(biplotprodrat))
biplotprodrat.row = apply(biplotprodrat.prop, 1, sum)
biplotprodrat.col = apply(biplotprodrat.prop, 2, sum)
rdiag = diag(1/ sqrt(biplotprodrat.row))
cdiag = diag (1/ sqrt (biplotprodrat.col))
biplotprodrat.row = as.matrix(biplotprodrat.row)
biplotprodrat.col = as.matrix(biplotprodrat.col)
E= rdiag %*% (biplotprodrat.prop - biplotprodrat.row %*%t (biplotprodrat.col)) %*% cdiag
biplotprodrat.sva = svd(E)
delta = biplotprodrat.sva$d
U= biplotprodrat.sva$u[,1:2]
V= biplotprodrat.sva$v[,1:2]

U[,1] = delta[1]* U[,1]/ sqrt (biplotprodrat.row)
U[,2] = delta[2]* U[,2]/ sqrt (biplotprodrat.row)
V[,1] = delta[1]* V[,1]/ sqrt (biplotprodrat.col)
V[,2] = delta[2]* V[,1]/ sqrt (biplotprodrat.col)

U = rbind(U,V)
inertia = sum(delta[delta^2])
per1 = delta[1]^2 / inertia
per2 = delta[2]^2 / inertia

dim1 = dim(biplotprodrat)[1]
ds =as.integer(dim1 +1)
dim2 = dim(biplotprodrat)[2]
dt = dim1 + dim2

plot(U[,1], U[,2], type= "n", xlab = paste("coord 1 - %inertia =  ", format(per1)),
     ylab = paste("coord 2 - % inertia = ", format(per2)))
text(U[1: dim1,1], U[1: dim1, 2], labels = dimnames(biplotprodrat)[[1]], col = 2)
text(U[ds : dt,1], U[ds: dt, 2], labels = dimnames(biplotprodrat)[[2]], col =4)
abline(h= 0, v=0, lty =2)
title ("biplot for indi and var")
```
















